---
title: "From staying time records to REST density estimates"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{From staying time records to REST density estimates}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = requireNamespace("ctrest", quietly = TRUE)
)
```

This vignette runs the whole workflow on simulated data: per-video staying
time records are merged into staying events with STwrapper, and animal density
is estimated with the REST model of
[ctrest](https://github.com/YoshihiroNakashima/ctrest). The last chunks need
ctrest:

```{r, eval = FALSE}
remotes::install_github("YoshihiroNakashima/ctrest")
```

## Simulate camera-trap videos

`simulate_stay_data()` simulates animals entering the focal area of 30
cameras over 60 days and annotates every recorded 20-second video. Animals
that stay longer than one video appear in consecutive videos, marked with
`Enter_cont = TRUE`.

```{r}
library(STwrapper)

set.seed(123)
sim <- simulate_stay_data(
  n_station = 30,
  days = 60,
  density = 30,      # individuals / km^2
  focal_area = 2,    # m^2
  stay_mean = 10,    # seconds
  stay_sd = 8,
  group_mean = 1.5,
  species = "deer"
)
truth <- attr(sim, "truth")
str(truth)
```

The records look like the annotation table of a real survey. Here the data
are written to a CSV file and read back, as you would do with your own data.

```{r}
path <- tempfile(fileext = ".csv")
write.csv(sim, path, row.names = FALSE)
stay_records <- read.csv(path)

head(stay_records[stay_records$Enter_cont %in% TRUE |
                    c(stay_records$Enter_cont[-1] %in% TRUE, FALSE),
                  c("video_name", "species", "enter", "out",
                    "stayingTimeCensoringType", "Enter_new", "Enter_cont")])
```

## Merge staying events with STwrapper

`convert_stay()` merges rows of the same staying event across videos and
returns the columns of `ctrest::detection_data`: one row per staying event
with `y = 1`, the staying time `Stay` (seconds) and the right-censoring
indicator `Cens`, plus rows with `y = 0` for videos in which the species was
detected without a new entry.

```{r}
detection <- convert_stay(stay_records, term = "term1")
head(detection)
table(y = detection$y, useNA = "ifany")
```

Every entry is counted once and has one staying time:

```{r}
c(entries = sum(detection$y, na.rm = TRUE),
  staying_times = sum(!is.na(detection$Stay)),
  simulated_entries = truth$n_entry)
```

## Estimate density with ctrest

The station table only needs station IDs here. The survey effort of each
camera is taken from its first and last videos (camera set-up and retrieval).

```{r, message = FALSE}
library(ctrest)

station_data <- data.frame(Station = unique(detection$Station))

station_rest <- format_station_data(
  detection_data = detection,
  station_data = station_data,
  col_name_station = "Station",
  col_name_species = "Species",
  col_name_y = "y",
  model = "REST"
)

station_effort <- add_effort(
  detection_data = detection,
  station_data_formatted = station_rest,
  col_name_station = "Station",
  col_name_datetime = "DateTime",
  col_name_term = "Term"
)

stay_data <- format_stay(
  detection_data = detection,
  station_data = station_data,
  col_name_station = "Station",
  col_name_species = "Species",
  col_name_stay = "Stay",
  col_name_cens = "Cens"
)

activity_data <- format_activity(
  detection_data = detection,
  col_name_station = "Station",
  col_name_species = "Species",
  col_name_datetime = "DateTime",
  indep_time = 30
)
```

```{r, message = FALSE, results = "hide"}
rest_model <- bayes_rest(
  formula_stay = Stay ~ 1,
  formula_density = ~ 1,
  formula_enter = ~ 1,
  station_effort_data = station_effort,
  stay_data = stay_data,
  activity_data = activity_data,
  activity_estimation = "kernel",
  stay_family = "lognormal",
  focal_area = truth$focal_area,
  cores = 2,
  iter = 4000,
  warmup = 1000,
  chains = 2,  # R CMD check allows at most 2 parallel processes
  thin = 2,
  model = "REST",
  target_species = "deer"
)
```

```{r}
rest_model$summary_result
```

The estimates can be compared with the values used in the simulation:

```{r}
c(density = truth$density, mean_stay = truth$mean_stay,
  activity_level = truth$activity_level)
```
